---
title: Test of Solarized Syntax Highlighting
subtitle: Here we go
author: TJ Mahr
# date: `r Sys.Date()`
output: 
  solarizeddocx::document: default
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = TRUE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

```{r}
# 75 characters fit okay
0123456789012345678901234567890123456789012345678901234567890123456789012345
```


Syntax gauntlet:

```{r, eval = FALSE}
library(magrittr)
x <- TRUE | T
y <- FALSE || F
1L + 1L
1.000 - 1.000
10 ^ 2
"string"
'string'
package::function
package::my_function
NA
NA_character_
Inf
NULL
1:10
for ( x in 1) {
  !! no
  f <- function(...) {
    x <- c(..., ..1, ..2, ...elt(1))
    return(data)
  }
  f <- \(x) {
    c(x)
  }
}

a <- list(x = 10, y = 100)
a$x
a[[1]]
a[2]

"newline\n and escaped slash \\"

x %>% rnorm(10, .) %>% round(2)

x |> round(2)
10.11 |> (\(x) round(x))()
-2 + 4i
```


Changes to the from the default language definition:

- Prompt comments `#>` and regular comments `#` get separate highlighting style

```{r}
# this should be italic but the output should be upright
1:10
```

- Explicit integers now highlighted

```{r}
x <- 1L
y <- 1
```

Gaps in the current syntax highlighting:

- It would be nice to highlight the package name and italicize it here:

```{r, eval = FALSE}
package::func
```

- Native pipe
- i in complex numbers


***


```{r, error = TRUE}
test %>% 
  testy() %>% 
  testing()
message("test")
warning("test")
stop("test")
# stop <- function(...) message(...)

if (TRUE) {
  print("hello")
} else {
  y ~ x
}
```

We also `have to test` markup styles

* list 1
* list 2
  - list 1
* list 3

> blockquote

```
print("hello")
```


https://bookdown.org/yihui/rmarkdown/new-formats.html
