---
title: Solarized Syntax Highlighting in Microsoft Word
subtitle: More than just highlighting scheme
author: TJ Mahr
date: |
  `r format(Sys.time(), "Updated on %A, %B %d, %Y %I:%M %p")`
output: 
  solarizeddocx::document: default
  word_document: default
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

## Overview

This package was created because I think the [solarized color
palette](https://github.com/altercation/solarized) looks nice in Word documents
created by pandoc, but solarized is not one of the supported highlighting
styles. 

This document is the result of calling `rmarkdown::word_document()` but
customizing the output by passing along three files to pandoc.

First, the **reference docx** file defines the styles for the Word document. The
main changes from the default style are that: 

0.  the "normal" text style now includes 10-pt of space before paragraphs. The
    "source code" blocks outputted by pandoc inherit from this style---I think
    this is how it works---so that space prevents code blocks from touching
    preceding text.
1.  the blue in headings and links uses solarized's blue color
2.  the `verbatim text` uses solarized's light body text color
3.  titles and other front matter are now left aligned
4.  the block text style is 1-pt smaller than the body text

```{r}
basename(solarizeddocx::file_reference_docx())
```

Second, the **solarized light highlighting theme** file defines which colors to assign
token types during syntax highlighting. 

```{r}
basename(solarizeddocx::file_solarized_light_theme())
```

I borrowed the color assignments mostly from [the rsthemes
package](https://github.com/gadenbuie/rsthemes/blob/main/inst/templates/solarized-light.scss).
These are:

```{scss, eval = FALSE}
// R language colors
$code_comment  : $base1;
$code_string   : $cyan;
$code_function : $blue;
$code_reserved : $green;
$code_operator : $yellow;
$code_bracket  : $yellow;
$code_value    : $magenta;
$code_namespace: $red;
$code_variable : $violet;
$code_message  : $orange;
$code_namespace_font_style: italic;

$code_other: $code_variable;
```

Third, I customized the **R syntax language definition file** so make results
somewhat better for Word documents. The most important change is that regular
comments are light and italic and prompt-comments used for output are darker and
more upright. This change makes it easier to read output from dataframes.

```{r}
# this comment is italic. the output below is roman
basename(solarizeddocx::file_solarized_light_theme())

model <- lm(mpg ~ cyl, mtcars)
summary(model)

coef(summary(model))
```

<!-- This inserts a page break in word. -->

```{=openxml}
<w:p>
  <w:r>
    <w:br w:type="page"/>
  </w:r>
</w:p>
```

## Code examples

Some basic examples:

```{r}
f <- function(x, ...) {
  mean(x, ...)
}

f(c(1, 2, 3, 4, NA), na.rm = TRUE)

print("Hello!")
```

Here is how wide we can make our code examples, comfortably, in Word.

```{r, eval = FALSE}
# 75 characters fit okay
0123456789012345678901234567890123456789012345678901234567890123456789012345
```

Here are the changes I made to the default R language definition:

```{r}
# italic code comments
print("but roman output comments")

# integers ending in L are now highlighted
c(1.0, 1, 1L)

# italicized package names in namespace calls
f <- stats::rnorm
f <- stats:::rnorm

# native pipe highlighted
x <- rnorm(10) |> round(1)
```

One feature we find in pandoc highlighting but do not find in RStudio is the
highlighting of escape sequences like `\n`. 

```{r}
# escape sequences are highlighted inside of strings
x <- "this string has an \"escaped\" characters"
y <- "we escape things by using a \\. \nso that \\n is a newline"
cat(x)
cat(y)
```

Based on my tests, two things that are currently not correct with the default
language definition and my modified language definition are the native pipe,
lambda syntax, and imaginary numbers.

```{r, eval = FALSE}
# lambda
round2 <- \(x) round(x, 2)
1 |> rnorm(1) |> round(2)

# immediately invoked lambda in a pipe
10.111 |> (\(x) round(x, 0))()

# imaginary number
-2 + 4i
```


### The Gauntlet

Below is the syntax gauntlet where we try a bunch of things.

```{r, eval = FALSE}
library(magrittr)

x
x2
x_2

# I don't think T and F should highlight because they are 
# overwritable
x <- TRUE | T
y <- FALSE || F

1L + 1L %% 2 == 0
1.0 * 1.0 / 1.0
+10 ^ 2 ** 2
-2 - 4i
.111
pi

package::my_function
package:::my_secret_function

NA
NA_character_
NA_complex_
NA_integer_
NA_real_
NaN
Inf
-Inf
NULL

for (x in 1:10) {
  !! no
  f <- function(...) {
    x <- c(..., ..1, ..2, ...elt(1))
    return(data)
  }
  f <- \(x) {
    c(x)
  }
  repeat
  break
  next
  while (FALSE) {
    print("never ever run"); print("not ever")
  }
}

# indexing
a <- list(x = 10, y = 100)
var <- "x"
a$x
a[1]
a["x"]
a[var]
a[[1]]
a[["x"]]
a[[var]]

# strings
'string'
"string"
"newline\n and escaped slash \\"

# names
`x y` <- "test"
`%in%`("a", letters)

# pipes
x %>% 
  rnorm(10, .) %>% 
  round(2)

x |> 
  round(2)

10.11 |> 
  (\(x) round(x))()

# I noticed message vs warning/stop having different RStudio colors
message("test")
warning("test")
stop("test")

if (TRUE) {
  print("hello")
} else {
  y ~ x + (1 | z:w)
}
```

Expanded string gauntlet from `?Quotes`

```{r}
# \n	newline
# \r	carriage return
# \t	tab
# \b	backspace
# \a	alert (bell)
# \f	form feed
# \v	vertical tab
# \\	backslash \
# \'	ASCII apostrophe '
# \"	ASCII quotation mark "
# \`	ASCII grave accent (backtick) `
# \nnn	character with given octal code (1, 2 or 3 digits)
# \xnn	character with given hex code (1 or 2 hex digits)
# \unnnn	Unicode character with given code (1--4 hex digits)
# or \u{nnnn} 
# \Unnnnnnnn	Unicode character with given code (1--8 hex digits)
# or \U{nnnnnnnn}

(hw2 <- "\110\x65\154\x6c\157\x20\127\x6f\162\x6c\144\x21")
(hw3 <- "\u48\u65\u6c\u6c\u6f\u20\u57\u6f\u72\u6c\u64\u21")
("\u0126\u0119\u1114\u022d\u2001\u03e2\u0954\u0f3f\u13d3\u147b\u203c")
("\U0126\U0119\U1114\U022d\U2001\U03e2\U0954\U0f3f\U13d3\U147b\U203c")

# 'Raw character constants are also available using a syntax similar to
# the one used in C++: r"(...)" with ... any character sequence, except
# that it must not contain the closing sequence )". The delimiter pairs
# [] and {} can also be used, and R can be used in place of r. For
# additional flexibility, a number of dashes can be placed between the
# opening quote and the opening delimiter, as long as the same number of
# dashes appear between the closing delimiter and the closing quote.'

r"(c:\Program files\R)"
r"{(\1\2)}"
r"(use both "double" and 'single' quotes)"
r"---(\1--)-)---"
```


## Markup styles

We also would like to test markup styles here.

***

# H1

## H2

### H3

#### H4

##### H5

###### H6


* list 1
* list 2
  - list 1
* list 3

Paragraph.

1. list 1
2. list 2
   1. list 1
3. list 1

> blockquote

A paragraph followed by code. **Strong**. *Emphasis*. `Code`.

```
print("hello")
```

Code that emits a table:

```{r}
knitr::kable(
  data.frame(
    day = c(1, 2, 3, 4),
    result = c("yes", "no", "yes", "yes")
  ), 
  caption = "Check it out"
)
```

